\bf \Large A Model Evaluation Study for Treatment Planning of Laser Induced Thermal Therapy

### 2015-04-21 Reviewer_response.tex

{\color{red}

Regarding the topic of what accuracy is required to consider a model useful
for prediction, it is a multifaceted concept. Firstly, we consider a model
prediction as useful in a particular case if the models ablation region and
the ablation as indicated via MR temperature imaging has a Dice similarity
coefficient (DSC) greater than 0.7. Ultimately the quantitative value of
DSC=.7 is based on prior information and experiences in image processing
literature~\cite{yung2010quantitative,Dice1945measures,zou2004three};  
in which DSC$>$.7 is considered `good agreement'. This point is highlighted
in the discussion section where the DSC is introduced.
 Secondly, the models overall accuracy in
a cohort is measured by the ratio of useful predictions versus inaccurate
predictions (\textit{i.e.}, number of DSC $\geq$ 0.7 versus DSC $<$ 0.7).
This overall accuracy is desired to be as high as possible, but it is beyond the
scope of our present investigation to name a success rate that would
indicate the model would be useful in clinical practice. This manuscript
focuses on reporting the accuracy achieved for our current computational
methods. Future work will consider the accuracy needed to impact treatment 
on a very carefully selected patient population.

}
\item
It is not clear how this method will improve pretreatment planning?

{\color{red}
Long term project goals are to develop a data driven computer model
that can prospectively predict the thermal dose outcome with high
accuracy. The ability to accurately predict the therapy outcome will enable
opportunities to optimize the delivery and maximized tumor kill while
minimizing healthy tissue damage. These efforts present results towards these goals
in the form of a retrospective analysis
with sufficiently large number datasets to evaluate prediction
statistics within a hypothesis testing framework. 
Our LOOCV analysis attempts to mimic our targeted prospective scenario
in that the model is trained on all previous (N-1) datasets and
the trained model prediction accuracy is evaluated on the
independent remaining dataset.
The success or failure of this paradigm is
to be related to several factors. 1) The model used must be able to
describe clinically relevant ablations. 2) There must be a sufficient
quantity of retrospective datasets for the
training to converge. 3) The cohort of retrospective datasets must have sufficient similarity within the
group and to the prediction scenario. \textit{E.g.}, a model trained on datasets from liver ablations is unlikely to
be predictive for brain. However, a training cohort of brain
metastases may or may not have enough homogeneity for a trained model
to be predictive. These points are added to the Discussion section.
}
\item
The introduction is too long. It is not clear how the second part that
describes the hardware relates to the paper.  There is no information
on the model size and the actual computations time.
It is not clear which hardware was used with which particular model.

{\color{red}
The hardware discussion was intended to
motivate the discussion on the tradeoff between algorithm efficiency and
accuracy on state of the art computing architectures. 
Similar to the previous comment, this diffuses the modeling accuracy focus of the 
manuscript and has been reorganized to a discussion point.
References to quantitatively evaluating computational efficiency as a paper
focus have been removed from the intro.

@@

%but takes considerably longer time
%to run. It is difficult to succinctly express the difference in
%computation times for at least two
%reasons. During optimization for a given patient,
%the FEM model and steady state model may have a different
%number of optimization realizations. Secondly, the FEM model is transient and has multiple time steps within
%one realization whereas the steady state model naturally has a single time step. Therefore, we
%consider the simplest means to summarize computation time is to report the timing of optimization and LOOCV.
%Optimization of the same dataset may involve differing number of realizations for the two models; LOOCV only
%has one forward prediction for each dataset. In optimization, the FEM model takes {\color{green}[Some estimate of time]}; during LOOCV, it took
%{\color{green}[Some estimate of time]}. The steady
%state models LOOCV takes 246.6 seconds or 11.2 seconds on average per dataset. This indicates that both models
%could be used before and during surgery. The Results
%and Discussion have been amended to describe this
%information; the new text is red.
}
\end{enumerate}

Reviewer 2:\\
Comments and suggestions to improve the manuscript:\\
\begin{itemize}
\item
Please expand on any heating induced changes in model parameters. Was blood perfusion held constant for the
entire treatment duration? Blood perfusion is well
known to change during ablative exposures (as the
authors note, this is sometimes exploited to confirm post-ablation heating with contrast-enhanced imaging).  Have the authors explored any techniques for
incorporating temperature/thermal damage dependencies for blood perfusion? Suggest adjusting nominal
perfusion level based on time-temperature history (see, for example, Schutt, Med Phys, 2008).

{\color{red}
Thank you for this suggestion. Model parameters change with temperature was
not considered in this study. 
Adding additional temperature dependet
properties would add one more degree of freedom to 'fit' the data. 
We were concerned with overfitting in this regard.
As the hyperthermia/ablation literature states in concert, tissue properties
are dependent on temperature and thermal damage state and these dependencies
are important to consider in order to accurately predict
outcomes. 
However, this time-dependent phenomenon is mitigated in our investigation via shorter ablation
times relative to other thermal modalities. \textit{E.g.}, laser ablations typically last 1.5 minutes to 2 minutes
whereas the RFA procedures modeled in Schutt, 2008, were 12 and 15 minutes in duration. Furthermore, and
perhaps more importantly, we do not consider the parameters from model optimization (or model
training) to be truly physical. In this investigation, the modeling is driven by model training, not model
complexity/physical fidelity. 
Discussion points for temperature dependent constitutive properties have
been updated.
}
\item
Did the authors investigate the impact of temperature dependent optical properties? During the model
calibration procedure, it appears a single optical attenuation coefficient was determined for all spatial
location/time points. Since optical absorption properties likely change during the course of heating,
one approach may be to perform the model calibration over only the first ~30-60 seconds, to estimate the
nominal optical absorption coefficient. This nominal value could then modified during heating based on
previously published reports of the relationship between optical attenuation and temperature. Suggest
considering approach where

### 2015-04-21 gpuPlanningAug13.tex
ENH: all review comments addressed

This manuscript focuses on the development of a practical methodology for
evaluating computer model predictions for \textit{a priori} planning the procedure  
given $N$ datasets from previous procedures.
Evaluation focuses on the trade-off between (1) algorithmic and numerical
implementation efficiency (2) against, most importantly, prediction accuracy.
Retrospective analysis of MR thermometry data acquired during previous procedures is
essential to train or calibrate the computer model parameters. 
The machine learning and statistics community have a rich history in
applying various algorithmic and physics-based data models to reach conclusions from a
given dataset~\cite{Breiman2001,Hastie2005}. 
Here we assume that a Pennes bioheat transfer model~\cite{Pennes1948} provides
representative predictions of the  physical process underlying the heating
observed within the MR thermometry data.
This physics-based approach
provides a theoretically sound and concise methodology to statistically summarize the high
dimensional thermometry dataset with a low dimensional model parameter subset.

The computer modeling approach targets a practical clinical setting of the operating room 
or surgical suite where portable computing hardware, efficient software implementations,
and efficient algorithm implementations are needed to deliver responsive predictions
of the treatment outcome.  Two distinct modeling approaches are pursued: 
(1) A GPU implementation of an unstructured hexahedral spectral element method
for predicting the bioheat transfer is developed. 
(2) A computationally inexpensive algorithm for predicting the heat transfer
from a linear superposition of analytic point source heating functions is also
presented as a reference implementation.

Combined with the data, the two modeling approaches presented provide an
environment to critically evaluate model selection for therapy planning of MRgLITT. 
%``Essentially, all models are wrong, but some are useful'', Box~\cite{box1987empirical},
Both approaches 
build intuition in the prediction by
repetitively training the underlying physics model 
to statistically match representative datasets. 
Predictions are critically evaluated in terms of solution efficiency and
accuracy for prospective treatment planning of MRgLITT procedures. 
Leave-One-Out Cross Validation (LOOCV) is used to simulate the clinical
scenario in which $N$ datasets from previous procedures are available to
calibrate the computer model. LOOCV provides an objective framework to
critically estimate the accuracy and confidence in predicting the outcome
of the procedure for the $N+1$ patient~\cite{Stone1974,Geisser1975,Kohavi1995,Browne2000,Arlot2010}.

@@

All computations were performed on the template hexahedral mesh
shown in Figure~\ref{LaserApplicatorOverview} (b).
For each simulation, the template was registered to the observed
laser location for each patient.
The mesh consists of disjoint regions for the applicator and tissue.
A quadrilateral mesh was extruded axially along the applicator to create
the base of the hexahedral finite element mesh. The mesh for the tissue
conforms to the surface of the application and extends sufficiently far
to ensure that the boundary does not influence the heating.
The discretization consists of $N_{dof}=$844,032 total GLL nodes. 
The degrees of freedom across the volume of the applicator were removed;
the effect of the room-temperature cooling fluid which protects the laser fiber
during heating were considered through the boundary conditions at the surface
nodes.

@@

{\color{red}For the two models, the time to run LOOCV
was recorded. The FEM model took ZYX for LOOCV, or asdfasdf on average per dataset.
The steady state took 247 seconds for LOOCV, or 11.2 seconds on average per dataset. }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{blue}
Previous real-time implementations of the kernel for
predicting the laser induced bioheat transfer were made feasible via
implementation on a massively parallel supercomputers~\cite{fuentesetal09,
Odenetal07,Dilleretal07,Fengetal07,Fuentesetal08}. 
Despite the increasing presence of large computational clusters within
large academic hospitals, reliance on such an approach makes clinical
translation of this research extremely difficult. This technology must
be practically available to any clinic using an MR-guided thermal therapy
system whereas integration with an institution's cluster and intranet creates
significant overhead and potential complications.
While modern single-chip multi-core CPU's have a peak performance of
$\mathcal{O}(10-100)$ GFLOPs and are sufficient 
for inexpensive algorithms such as the linear superposition of analytic functions presented
in Section~\ref{AnalyticBioheatModel}, 
current GPUs~\cite{Farber2011a,Muller2013} have a peak
performance of $\mathcal{O}(1-2)$ TFLOPs (single precision) and the
portability of these devices are ideal for developing more advanced 
numerical discretizations of the underlying partial differential equations
for guiding the therapy. However, a significant
time investment is needed to express the class of memory bound numerical
methods used in simulation within an algorithmic framework that can
simultaneously exploit these floating point capabilities and minimize the
global memory access latency of these devices.
% 28 = 1030/ (148/4) single precision.
% max floating point operations per second is 1030 GFlop
% data transfer of single precision numbers to the cores = bandwidth x data type size = 148  x (4bytes / float)
% >>> 1030./ (148./4.)
% 27.83783783783784
For example, the $\mathcal{O} (150)$~GB/s memory bandwidth of the Tesla M2070
used in the manuscript requires $\approx$ 30 floating point operations per
single precision global memory access to fully utilize the floating point capabilities
of the GPU. 
The spectral element method presented in Section~\ref{BioheatComputerModel}
achieves efficient global memory access through a matrix-free
implementation of the inherent linear system of equations.
As opposed to explicitly storing and reading matrix entries
from global memory, the matrix-free method recalculates the local matrix entries
needed within the linear system solve.
This memory access design pattern has demonstrated a 4-10x speedup  over
the matrix explicit methods~\cite{Muller2013,Knepley_2_2013,Medina2014}.
}

{\color{red}We are presenting this manuscript to introduce the possibility
of a new paradigm of treatment planning via a train-and-predict scheme.
This method is becoming possible because of a growing number of retrospective patient
datasets. Relatively simple models were chosen for this demonstration
to expedite the model training, \textit{i.e.}, optimization. The success
or failure of this paradigm is expected to be related to several factors.
First, there must be a sufficient quantity of retrospective datasets
for the training to converge. Second, the cohort of retrospective
datasets used in training must have sufficient similarity within the group
and to the prediction scenarios. Third, the model used must be able 
to describe clinically relevant ablations. In this investigation,
none of these three factors are satisfied. Nonetheless,
we propose that model training may expedite clinically useful,
patient-specific prediction by reducing the model complexity necessary
and/or the need for patient-specific or population-based parameter
measurements. }

{\color{red}Given the number of available datasets, the LOOCV analysis
provides the best available recapitulation of the clinical scenario}
for treatment planning. Prior knowledge and experience is embodied
within the thermometry data from the previous $N$ ablations.
Information from the previous ablations is extracted by calibrating
a computer model to the available data. 
In this case, we calibrated our optical parameter to the thermometry data.

@@

The two models presented provides a canonical model selection comparison of the
trade-off between the time investment in the algorithm, efficiency of the
numerical implementation, and accuracy required for predicting the final endpoint
of the application. The two-sided, two-sample paired Wilcoxon signed rank test
indicates the predictive results' medians are significantly different and
only the transient model's median DSC performance was significantly
$\geq$ 0.7. 
The finite element discretization of the governing equations requires significant
expertise of the GPU computing architecture as well as detailed algorithmic
understanding of both the finite element technology and matrix-vector multiply
within the iterative linear system solver to maximize floating point operation
throughput and minimizes memory transactions latencies.
Meanwhile, the steady state superposition analysis, Equation~\eqn{GreenSuperPosition},
provides treatment predictions with fewer floating point operations and would be selected under an Occam
razor~\cite{jaynes2003probability} philosophy highly preferential to simplicity.
While all kernels
for the present spectral element methods and preconditioned conjugate gradient
method were hand coded in this manuscript, library implementations of the
matrix-free iterative solver approach are also
appearing~\cite{Knepley2013,Knepley2014}.

%{\color{red}The focus of this manuscript is to compare two models' accuracy
%in predicting the maximum ablation size. However, for any given predictive model,
%a natural question to address is how much accuracy is necessary to clinically useful?
%A second question to answer is 
%The

The finite element discretization of the governing equations, however, 
provides significant opportunity for further
physics-based improvements including 
higher order model spherical harmonic expansions in the laser fluence 
model~\cite{carp2004radiative}. 
As seen in Figure~\ref{QuantitativeComparison}, the models used
introduce a bias in the model parameter
recovery that differs from published literature values.
{\color{red}The bias may arise from inaccuracies in the modeling assumptions, perhaps
most significant being the use of optical tissue properties that are
invariant in space and time. The literature has clear examples
where temperature/damage dependent and spatially/temporally dependent
parameters are critical to the
prediction~\cite{Mohammed05,Schutt2008}.
The choice to use a single constant parameter for every
optimized MRTI dataset is driven by
several considerations. The focus of this investigation
is to demonstrate the use of training in the two models.
The inclusion of spatially varied, damage dependent, and
multiple parameters would require considerably more
complicated computational methods and resources; this
should be persued in the future. Higher order physics
models of the fluence are be expected to provide the highest accuracy in
recovering the optical parameters during the calibration process and for
characterizing the tissue properties. }

The finite element discretization of the governing equations
also provides a rigorous physics-based methodology to 
incorporate tissue heterogeneities into the treatment prediction.
Calibrations of the spatially heterogeneous 
optical parameter field has been shown to provide highly accurate 
predictions~\cite{fuentes2013inverse}.
Adjoint-based methods of computing the gradient of the objective 
function are necessary to efficiently optimize in the higher dimensional parameter space.
Tissue heterogeneities could similarly be incorporated
into the steady state superposition analysis, Equation~\eqn{GreenSuperPosition}, in a patient-specific
`\textit{ad hoc}' manner. However, this would violate the underlying homogeneous
tissue parameter assumptions which provided the mathematical
structure for the concise analytical solution.
A Gaussian process~\cite{rasmussen2006gaussian,Constantinescu2013} 
framework may be appropriate
in which the model parameters recovered may be interpreted as hyperparameters
for the covariance kernels.
Similar to physics-based model calibration, optimization of the
hyperparameters in the Gaussian process kernels offers a complementary
trade-off between data fitting and smoothing. Further, Gaussian processes allow
for prior information to be used and provide a full probabilistic prediction
and an estimate of the uncertainty.

@@

\paragraph{Conclusion}
Currently, the neurosurgeon reviews anatomical MR images to plan his/her
trajectory to reach the tumor with neuro-navigation software, but does not
have the capability to visualize outcomes of the laser ablation using
various trajectories beforehand. 
Fully developed and commercially implemented predictive computer models 
will extend this functionality to include \textit{a priori} visualization of the
potential outcomes for complex treatment scenarios (multiple
applicators/trajectories) in which the laser ablation is performed near a
critical structure within a heterogeneous tissue environment. 
This work presents a step in this direction and demonstrates
the feasibility in establishing a confidence in these predictions.
Consideration of other available metadata to improve prediction accuracy
is a topic of on-going research.

>>>>

This manuscript focuses on the development of a practical methodology for
evaluating computer model predictions for \textit{a priori} planning the procedure  
given $N$ datasets from previous procedures.
%Evaluation focuses on the trade-off between (1) algorithmic and numerical
%implementation efficiency (2) against, most importantly, prediction accuracy.
{\color{red} Evaluation focuses on prediction accuracy.}
Retrospective analysis of MR thermometry data acquired during previous procedures is
essential to train or calibrate the computer model parameters. 
The machine learning and statistics community have a rich history in
applying various algorithmic and physics-based data models to reach conclusions from a
given dataset~\cite{Breiman2001,Hastie2005}. 
Here we assume that a Pennes bioheat transfer model~\cite{Pennes1948} provides
representative predictions of the  physical process underlying the heating
observed within the MR thermometry data.
This physics-based approach
provides a theoretically sound and concise methodology to statistically summarize the high
dimensional thermometry dataset with a low dimensional model parameter subset.

%The computer modeling approach targets a practical clinical setting of the operating room 
%or surgical suite where portable computing hardware, efficient software implementations,
%and efficient algorithm implementations are needed to deliver responsive predictions
%of the treatment outcome.  
Two distinct modeling approaches are pursued: 
(1) A GPU implementation of an unstructured hexahedral spectral element method
for predicting the bioheat transfer is developed. 
(2) A computationally inexpensive algorithm for predicting the heat transfer
from a linear superposition of analytic point source heating functions is also
presented as a reference implementation.
{\color{red}
Combined with the data, the two modeling approaches presented provide an
environment to critically evaluate model accuracy and  selection for therapy
planning of MRgLITT. 
}
%``Essentially, all models are wrong, but some are useful'', Box~\cite{box1987empirical},
Both approaches 
build intuition in the prediction by
repetitively training the underlying physics model 
to statistically match representative datasets. 
Predictions are critically evaluated in terms of solution efficiency and
accuracy for prospective treatment planning of MRgLITT procedures. 
Leave-One-Out Cross Validation (LOOCV) is used to simulate the clinical
scenario in which $N$ datasets from previous procedures are available to
calibrate the computer model. LOOCV provides an objective framework to
critically estimate the accuracy and confidence in predicting the outcome
of the procedure for the $N+1$ patient~\cite{Stone1974,Geisser1975,Kohavi1995,Browne2000,Arlot2010}.

@@

All computations were performed on the template hexahedral mesh
shown in Figure~\ref{LaserApplicatorOverview} (b).
For each simulation, the template was registered to the observed
laser location for each patient.
The mesh consists of disjoint regions for the applicator and tissue.
A quadrilateral mesh was extruded axially along the applicator to create
the base of the hexahedral finite element mesh. The mesh for the tissue
conforms to the surface of the application and extends sufficiently far
to ensure that the boundary does not influence the heating.
The discretization consists of $N_{dof}=$844,032 total GLL nodes. 
The degrees of freedom across the volume of the applicator were removed;
the effect of the room-temperature cooling fluid which protects the laser fiber
during heating were considered through the boundary conditions at the surface
nodes. 
{\color{red}
A mesh resolution of approximately the 1mm pixel size near the applicator was used
to ensure convergence of the discretized solution~\cite{fuentesetal11a}. 
Multiple time resolutions were also evaluated to ensure convergence of the
time stepping scheme.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given the number of available datasets, the LOOCV analysis
provides a methodology to recapitulate the clinical scenario
for treatment planning. Prior knowledge and experience is embodied
within the thermometry data from the previous $N$ ablations.
Information from the previous ablations is extracted by calibrating
a computer model to the available data. 
In this case, we calibrated our optical parameter to the thermometry data.
{\color{red}
%We are presenting this manuscript to introduce the possibility of a new
%paradigm of treatment planning via a train-and-predict scheme.  This method
%is becoming possible because of a growing number of retrospective patient
%datasets. 
%% Relatively simple models were chosen for this demonstration
%% to expedite the model training, \textit{i.e.}, optimization. 
Modeling goals are to ultimately utilize the calibrated model in optimizing
the thermal dose delivery.
The success
or failure of this paradigm is related to several factors.
First, there must be a sufficient quantity of retrospective datasets
for the training to converge. Second, the cohort of retrospective
datasets used in training must have sufficient similarity within the group
and to the prediction scenarios. Third, the model must be able 
to describe clinically relevant ablations. 
%In this investigation, none of these three factors are satisfied. 
%Nonetheless,
%we propose that model training may expedite clinically useful,
%patient-specific prediction by reducing the model complexity necessary
%and/or the need for patient-specific or population-based parameter
%measurements.

}

@@

The two models presented provides a canonical model selection comparison of the
trade-off between the time investment in the algorithm, efficiency of the
numerical implementation, and accuracy required for predicting the final endpoint
of the application. The two-sided, two-sample paired Wilcoxon signed rank test
indicates the predictive results' medians are significantly different and
only the transient model's median DSC performance was significantly
$\geq$ 0.7. 
{\color{red}
As a quantitative reference,
the forward solve mapping between model parameters, Table~\ref{modeldata}, 
and temperature field is considered the fundamental computational
operation of this study.
Total runtimes in the analysis are  proportional to the total number of foward solve
iterations in the optimation and LOOCV analysis. 
The run times for the forward solves in this study 
averaged 187.e3s and 11.2s for the transient model GPU implementation 
and steady state CPU implementation; respectively. 
These numbers are intended to provide intuition for the observed
practical run times on a local CPU 
% (Intel Xeon, 6 core, 2.4 GHz)  
(Intel Xeon, 6 core, 2.4 GHz, double precision peak 57 Gflops )  
workstation with attach GPU accelerators
%(NVIDIA Tesla M2070)
(NVIDIA Tesla M2070 double precision peak 515 Gflops) 
available to this study.
}
The finite element discretization of the governing equations requires significant
expertise of the GPU computing architecture as well as detailed algorithmic
understanding of both the finite element technology and matrix-vector multiply
within the iterative linear system solver to maximize floating point operation
throughput and minimizes memory transactions latencies.
As opposed to explicitly storing and reading matrix entries
from global memory, the matrix-free method recalculates the local matrix entries
needed within the linear system solve.
This memory access design pattern has demonstrated a 4-10x speedup  over
the matrix explicit methods~\cite{Muller2013,Knepley_2_2013,Medina2014}.
Meanwhile, the steady state superposition analysis, Equation~\eqn{GreenSuperPosition},
provides treatment predictions with fewer floating point operations and would be selected under an Occam
razor~\cite{jaynes2003probability} philosophy highly preferential to simplicity.
While all kernels
for the present spectral element methods and preconditioned conjugate gradient
method were hand coded in this manuscript, library implementations of the
matrix-free iterative solver approach are also
appearing~\cite{Knepley2013,Knepley2014}.

%% {\color{red}
%% 
%% Previous real-time implementations of the kernel for
%% predicting the laser induced bioheat transfer were made feasible via
%% implementation on a massively parallel supercomputers~\cite{fuentesetal09,
%% Odenetal07,Dilleretal07,Fengetal07,Fuentesetal08}. 
%% Despite the increasing presence of large computational clusters within
%% large academic hospitals, reliance on such an approach makes clinical
%% translation of this research extremely difficult. This technology must
%% be practically available to any clinic using an MR-guided thermal therapy
%% system whereas integration with an institution's cluster and intranet creates
%% significant overhead and potential complications.
%% While modern single-chip multi-core CPU's have a peak performance of
%% $\mathcal{O}(10-100)$ GFLOPs and are sufficient 
%% for inexpensive algorithms such as the linear superposition of analytic functions presented
%% in Section~\ref{AnalyticBioheatModel}, 
%% current GPUs~\cite{Farber2011a,Muller2013} have a peak
%% performance of $\mathcal{O}(1-2)$ TFLOPs (single precision) and the
%% portability of these devices are ideal for developing more advanced 
%% numerical discretizations of the underlying partial differential equations
%% for guiding the therapy. However, a significant
%% time investment is needed to express the class of memory bound numerical
%% methods used in simulation within an algorithmic framework that can
%% simultaneously exploit these floating point capabilities and minimize the
%% global memory access latency of these devices.
%% % 28 = 1030/ (148/4) single precision.
%% % max floating point operations per second is 1030 GFlop
%% % data transfer of single precision numbers to the cores = bandwidth x data type size = 148  x (4bytes / float)
%% % >>> 1030./ (148./4.)
%% % 27.83783783783784
%% For example, the $\mathcal{O} (150)$~GB/s memory bandwidth of the Tesla M2070
%% used in the manuscript requires $\approx$ 30 floating point operations per
%% single precision global memory access to fully utilize the floating point capabilities
%% of the GPU. 
%% The spectral element method presented in Section~\ref{BioheatComputerModel}
%% achieves efficient global memory access through a matrix-free
%% implementation of the inherent linear system of equations.
%% 
%% }
%{\color{red}The focus of this manuscript is to compare two models' accuracy
%in predicting the maximum ablation size. However, for any given predictive model,
%a natural question to address is how much accuracy is necessary to clinically useful?
%A second question to answer is 
%The

The finite element discretization of the governing equations, however, 
provides significant opportunity for further
physics-based improvements including 
higher order model spherical harmonic expansions in the laser fluence 
model~\cite{carp2004radiative}. 
As seen in Figure~\ref{QuantitativeComparison}, the models used
introduce a bias in the model parameter
recovery that differs from published literature values.
{\color{red}The bias may arise from inaccuracies in the modeling assumptions, perhaps
most significant being the use of optical tissue properties that are
invariant in space and time. The literature has clear examples
where temperature/damage dependent and spatially/temporally dependent
parameters are critical to the
prediction~\cite{Mohammed05,Schutt2008}. The use of thermometry data during
only the test pulse is also expected to influence the recovered optical
properties.
The choice to use a single constant parameter for full time history of each
MRTI dataset is motivated by
the practical model training focus of this investigation.
Higher order physics
models of the fluence are be expected to provide the highest accuracy in
recovering the optical parameters during the calibration process and for
characterizing the tissue properties.

%Higher order methods in which constitutive nonlinearites and different
%subsets of training sets such as the patient specifi
%test pulse that are optimal for planning
%is considered future work. This requires overfitting studies, harmonization
%of meta data and studies on how much training data is needed to achieve
%the desired accuracy.
 }

The finite element discretization of the governing equations
also provides a rigorous physics-based methodology to 
incorporate tissue heterogeneities into the treatment prediction.
Calibrations of the spatially heterogeneous 
optical parameter field has been shown to provide highly accurate 
predictions~\cite{fuentes2013inverse}.
{\color{red} The inclusion of spatially varied, damage dependent, and
multiple parameters should be pursued in future efforts.}
Adjoint-based methods of computing the gradient of the objective 
function are necessary to efficiently optimize in the higher dimensional parameter space.
Tissue heterogeneities could similarly be incorporated
into the steady state superposition analysis, Equation~\eqn{GreenSuperPosition}, in a patient-specific
`\textit{ad hoc}' manner. However, this would violate the underlying homogeneous
tissue parameter assumptions which provided the mathematical
structure for the concise analytical solution.
A Gaussian process~\cite{rasmussen2006gaussian,Constantinescu2013} 
framework may be appropriate
in which the model parameters recovered may be interpreted as hyperparameters
for the covariance kernels.
Similar to physics-based model calibration, optimization of the
hyperparameters in the Gaussian process kernels offers a complementary
trade-off between data fitting and smoothing. Further, Gaussian processes allow
for prior information to be used and provide a full probabilistic prediction
and an estimate of the uncertainty.

@@

\paragraph{Conclusion}
Currently, the neurosurgeon reviews anatomical MR images to plan his/her
trajectory to reach the tumor with neuro-navigation software, but does not
have the capability to visualize outcomes of the laser ablation using
various trajectories beforehand. 
Fully developed and commercially implemented predictive computer models 
will extend this functionality to include \textit{a priori} visualization 
{\color{red} and optimization} of the
potential outcomes for complex treatment scenarios (multiple
applicators/trajectories) in which the laser ablation is performed near a
critical structure within a heterogeneous tissue environment. 
This work presents a step in this direction and demonstrates
the feasibility in establishing a confidence in these predictions.
Consideration of other available metadata to improve prediction accuracy
is a topic of on-going research.
